---
title: 'Load and Tidy data'
date: "Oct 30, 2021"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
    number_sections: true
---

```{r echo = FALSE}
# Course: OMSBA 5305 
# Purpose: Data tidying
# Date: Oct 30, 2021
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
```

+ Load packages 
```{r warning = FALSE, message = FALSE}
# Load packages
library(tidyverse) 
library(vtable)
library(readxl)     # to read excel file
library(dynlm)      # to read linear model
library(latticeExtra)
library(scales)
```

+ Load data
```{r}
oecd <- read_csv("../data/oecd.csv")
head(oecd)
```

+ Look at data
```{r}
vtable(oecd)
```

  - there are two different measurements: HCPCARSA and HVPVOBARSA
  
```{r}
unique(oecd$MEASURE) 
```

Per the documentation for this data set - (https://stats.oecd.org/fileview2.aspx?IDFile=a7e60ef1-5139-4a2c-9331-cfd135033b8a)
  - HCPCARSA: US dollars per head, at current prices and current PPPs, annual levels, s.a.
  - HVPVOBARSA: US dollars per head, at 2015 price level and 2015 PPPs, annual levels, s.a.
  
I'm assuming we'll want HVPVOBARSA, but I will see if I can confirm that is the variable we want to use.

 + Data tidying
```{r}
quarterly_oecd <- oecd %>% 
  filter(FREQUENCY == "Q", MEASURE == "HVPVOBARSA") %>%
  separate(Period, into = c("Quarter", "Year"), sep = "-") %>%
  mutate(Quarter = factor(Quarter), Year = factor(Year)) %>%
  select(Country, Quarter, Year, Unit, Value) 

vtable(quarterly_oecd)
```


 + Find 5 countries with highest GDP in 2019
```{r}
quarterly_oecd %>% 
  filter(Year == 2019) %>%
  group_by(Country) %>%
  summarize(avg_annual_gdp = (sum(Value)/4)) %>%
  arrange(-avg_annual_gdp)
```

- top5: Luxembourg, Ireland, Switzerland, Norway, and United States

+ check data for top5 

  - mainly I am looking to see what time periods each country has data for in this dataset
  
```{r}
## Luxembourg
Luxembourg_quarterly <- quarterly_oecd %>%
  filter(Country == "Luxembourg") 

Luxembourg_quarterly %>% group_by(Year) %>% count()
```

  - Luxembourg data starts in Q1 of 1995 through Q1 of 2021

```{r}
## Ireland
Ireland_quarterly <- quarterly_oecd %>%
  filter(Country == "Ireland") 

Ireland_quarterly %>% group_by(Year) %>% count()
```

- Ireland data starts in Q1 of 1995 through Q2 of 2021

```{r}
## Switzerland
Switzerland_quarterly <- quarterly_oecd %>%
  filter(Country == "Switzerland") 

Switzerland_quarterly %>% group_by(Year) %>% count()
```

  - Switzerland data starts in Q1 of 1991 through Q1 of 2021

```{r}
## Norway
Norway_quarterly <- quarterly_oecd %>%
  filter(Country == "Norway") 

Norway_quarterly %>% group_by(Year) %>% count()
```

  - Norway data starts in Q1 of 1995 through Q2 of 2021

```{r}
US_quarterly <- quarterly_oecd %>%
  filter(Country == "United States") 

US_quarterly %>% group_by(Year) %>% count()
```

  - US data starts in Q1 of 1990 through Q3 of 2021

```{r}
top_5_1995_to_2020 <- quarterly_oecd %>%
  filter(Country %in% c("Luxembourg", "Ireland", "Switzerland", "Norway", "United States"),
         Year %in% c(1995:2020)) 

q1_2021 <- quarterly_oecd %>%
  filter(Country %in% c("Luxembourg", "Ireland", "Switzerland", "Norway", "United States"), 
         Year == 2021, Quarter == "Q1")
         
df <- bind_rows(top_5_1995_to_2020, q1_2021)

df <- df %>%
  mutate(Country = factor(Country),
         Unit = factor(Unit)) %>%
  group_by(Country) %>%
  arrange(Year)
```

```{r}
vtable(df)
```

 - each country now has 105 observations

## EDA
### Non-Graphical EDA

```{r}
head(df)
```


```{r}
# factor variables
table(df$Value)
table(df$Quarter)
df %>% group_by(Country, Quarter) %>% summarise(sum(Value))
```

### Graphical EDA

```{r, fig.width=12, fig.height = 10}
df %>% ggplot(mapping = aes(x = Country, y = Value, color = Country)) +
              geom_boxplot()+
              labs(title = "Variations of GDP by Country",
                   subtitle = "Luxembourg has the hieghst GDP",
                   x = "Country", y = "Values by USD",
                   caption = "Source : OECD.org", tag = "Figure 1") +
             scale_y_continuous(labels = comma, breaks = seq(10000, 150000, 10000))
```

```{r}
df %>% ggplot(mapping = aes(x = Year, y = Value, color = Country)) +
              geom_point() + 
              geom_smooth()
```


```{r}
df %>% ggplot(mapping = aes(x = Year, y = Value, group = Country, color = Country)) +
              geom_line() +
              geom_point()
```

```{r, fig.width=12, fig.height = 8}
df %>% group_by(Country, Year) %>% summarise(Value = mean(Value)) %>%
       ggplot(mapping = aes(x = Year, y = Value, group = Country, color = Country)) +
              geom_line() +
              geom_point() +
              labs(title = "Variations of GDP per Year by Country",
                   subtitle = "Ireland has the most variation in GDP",
                   x = "Years", y = "Values by USD",
                   caption = "Source : OECD.org", tag = "Figure 2") +
             scale_y_continuous(labels = comma, breaks = seq(10000, 150000, 10000))
```

```{r}

df %>% ggplot(mapping = aes(x = Year, y = Value, color = Country)) +
  geom_point() + 
  geom_smooth()

df %>% ggplot(mapping = aes(x = Year, y = Value, group = Country, color = Country)) +
  geom_line() +
  geom_point()


top_5_full <- df %>% group_by(Country, Year) %>% summarise(Value = mean(Value)) %>%
  ggplot(mapping = aes(x = Year, y = Value, group = Country, color = Country)) +
  geom_line() +
  geom_point() +
  labs(title = "Variations of GDP per Year by Country",
       subtitle = "Ireland has the most variation in GDP",
       x = "Years", y = "Values by USD",
       caption = "Source : OECD.org", tag = "Figure 2") +
  scale_y_continuous(breaks = seq(10000, 150000, 10000))

ggsave(filename = "top_5_full.png", plot = top_5_full)

```

# final data set

```{r}
df <-  df %>% 
  filter(Country == 'Ireland')
df1 <- ts(df$Value, frequency=4, start = 1995)
```

